---
title: "RiboMethSeq — Quality control report"
subtitle: "Read end counts at all genomic positions"
date: "`r format(Sys.time(), '%x | %X %Z')`"
output:
  html_document:
    theme:
      version: 5
      bootswatch: flatly
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
    toc_depth: 2
    fig_width: 9
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(knitr)

# Reference order = order of appearance in the count files (= FASTA / bowtie
# index order). This keeps rRNA first and snRNA (U*) last, instead of the
# previous size-based sort which scrambled the list.
rna_names  <- fcount_data[[1]][[1]]
ref_levels <- unique(rna_names)
reference_rna_tbl <- as.data.frame(table(factor(rna_names, levels = ref_levels)))
colnames(reference_rna_tbl) <- c("Reference", "Positions")
reference_rna_tbl$Type <- ifelse(grepl("snRNA", reference_rna_tbl$Reference), "snRNA", "rRNA")
reference_rna_tbl <- reference_rna_tbl[, c("Reference", "Type", "Positions")]

n_samples   <- ncol(fcount_matrix)
n_positions <- nrow(fcount_matrix)
n_refs      <- nrow(reference_rna_tbl)
```

<style>
/* --- Headings: tone down the default flatly sizes, keep h1>h2>h3 --- */
h1.title    { font-size: 1.9rem !important; font-weight: 600; }
.subtitle   { font-size: 1.05rem !important; color: #6b7682; }
h1:not(.title) {
  font-size: 1.55rem !important;
  color: #2c7fb8 !important;
  border-bottom: 1px solid #e4e7eb;
  padding-bottom: 0.25rem;
  margin-top: 1.6rem;
}
h2          { font-size: 1.25rem !important; }
h3          { font-size: 1.05rem !important; }

/* --- Callouts: softer, lighter than default bootstrap alerts --- */
.alert {
  border-radius: 6px;
  padding: 0.7rem 1rem;
  font-size: 0.95rem;
  box-shadow: none;
}
.alert-info {
  background-color: #eef4fa !important;
  border: 1px solid #d4e2f0 !important;
  border-left: 4px solid #6aa0d8 !important;
  color: #2f3e4e !important;
}
.alert-warning {
  background-color: #fbf6ec !important;
  border: 1px solid #ecdfc6 !important;
  border-left: 4px solid #d6a93b !important;
  color: #5a4a22 !important;
}

/* --- Tables: light, thin lines, a bit of body --- */
table {
  border-collapse: collapse !important;
  margin: 1.1rem 0;
  font-size: 0.93rem;
  width: auto;
}
table caption {
  caption-side: top;
  color: #6b7682;
  font-style: italic;
  padding-bottom: 0.3rem;
}
table th, table td {
  padding: 0.45rem 0.9rem !important;
  border: 1px solid #e4e7eb !important;
}
table thead th {
  background-color: #f4f6f8 !important;
  border-bottom: 2px solid #dde1e6 !important;
  font-weight: 600;
  text-align: left;
}
table tbody tr:nth-child(even) { background-color: #fafbfc !important; }
table tbody tr:hover { background-color: #eef2f6 !important; }
</style>

# General information

```{r general-info}
kable(
  data.frame(
    Metric = c("Samples", "Reference RNAs", "Total positions"),
    Value  = c(n_samples, n_refs, n_positions)
  ),
  align = c("l", "r")
)
```

<div class="alert alert-info" role="alert">
Each reference RNA is analysed across **all** its genomic positions. The table
below lists the references included in this run, grouped by type
(**rRNA** first, **snRNA** last).
</div>

```{r reference-table}
kable(reference_rna_tbl, row.names = FALSE, align = c("l", "l", "r"))
```

# Coverage

The coverage corresponds to the number of end read counts at each individual genomic position.

<div class="alert alert-info" role="alert">
A minimum coverage of **100 end read counts** is required to reach a robust
C-score. The boxplot below shows the distribution of the absolute number of end
read counts per genomic position, for each sample. The blue line marks the
100-count threshold — a **median below it flags a putative outlier sample**.
</div>

```{r boxplot}
plot(plot_boxplot_samples(fcount_matrix,show_outlier=TRUE))
```

<div class="alert alert-info" role="alert">
C-score robustness also depends on overall coverage, so a sample with markedly
lower or higher coverage may be an outlier. The **Relative Log Coverage** plot
below shows each sample's variation around the series median; the blue lines are
median ± 2·MAD. A **median outside these lines flags a putative outlier**.
</div>

```{r RLE}
plot_RLE(fcount_matrix,"count",show_outlier = T)
```

# Correlation-based distance heatmap {.tabset}

<div class="alert alert-info" role="alert">
Coverage profiles are compared two-by-two across the whole series using Pearson
correlation. The distance heatmap summarises these scores, either **clusterized**
or **non-clusterized** (tabs below). A perfect correlation reads **0 (white)**,
no correlation **1 (red)**.
</div>

## Clusterized {.unnumbered}

```{r heatmap_corr, results = 'asis',fig.height=7}
tryCatch({
    plot_heatmap_corr(fcount_matrix)
}, error = function(cond) {
    cat("<div class=\"alert alert-warning\" role=\"alert\">Not enough samples to clusterize!</div>")
})
```

## non-clusterized {.unnumbered}

```{r heatmap_corr_tri, fig.height=7}
plot_heatmap_corr(fcount_matrix, use_triangle=TRUE)
```


# Coverage profile: series comparison

<div class="alert alert-info" role="alert">
The **Correspondence Analysis (COA)** below illustrates the distance between
samples based on their coverage profiles. **Isolated samples** can be considered
putative outliers.
</div>

```{r}
plot_coa(fcount_matrix)
```

<hr>
<p class="text-muted"><small>Automatically generated with RiboMethSeq-nf.</small></p>
