---
title: "Quality control using read end counts at all genomic positions"
output: html_document
---

```{r, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning=FALSE)

reference_rna_tbl <- as.data.frame(table(fcount_data[[1]][[1]]))
reference_rna_tbl <- reference_rna_tbl[order(as.numeric(reference_rna_tbl[[2]])),]

```

# General informations

* Number of samples : **`r ncol(fcount_matrix) - 1`**
* Number of positions : **`r nrow(fcount_matrix)`**
* Reference RNA(s) used : `r paste0("\n\t* **", reference_rna_tbl[[1]], "** with **", reference_rna_tbl[[2]], "** positions",collapse="")`

# Coverage

The coverage corresponds to the number of end read counts at each individual genomic position.

First, by comparing the C-score at a given genomic position depending on the number of end read counts in its local environment, it appears that a minimum coverage of 100 end read counts is required to reach a robust C-score. **Here is a boxplot given the distribution of the absolute number of end read counts at each genomic positions for all individual samples.** The blue line gives the minimal 100 end read counts. Median below the blue line identifies putative outlier sample.

```{r boxplot}
plot(plot_boxplot_samples(fcount_matrix,"count","log10(count)",show_outlier=TRUE))
```

Second, sample that presents a lower or higher coverage might be considered as outlier since C-score robustness is dependent on the coverage. **Here is a Relative Log Coverage plot given the variation of the end read counts to the serie/sample (?) median for each individual sample.** The blue lines give median +/- 2mad variation. Median outside the blue lines identifies putative outlier sample.

```{r RLE}
plot_RLE(fcount_matrix,"count",show_outlier = T)
```

# Correlation-based distance heatmap {.tabset}

First, to help in identifying distinct coverage profile, coverage profiles are compared between two samples in the whole series. Two-by-two samples pearson correlation of the coverage profile is calculated. **Here is given a distance heatmap that summarizes of the correlation scores in the entire series either as clusterized or non-clusterized (tab)**. A perfect correlation is given by 0 (white) and no correlation by 1 (red).

## Clusterized {.unnumbered}

```{r heatmap_corr, results = 'asis',fig.height=7}
tryCatch({
    plot_heatmap_corr(fcount_matrix)
}, error = function(cond) {
    cat("<span style=\"color: red;\">Not enough samples to clusterize !</span>")
})
```

## non-clusterized {.unnumbered}

```{r heatmap_corr_tri, fig.height=7}
plot_heatmap_corr(fcount_matrix, use_triangle=TRUE)
```

Second, to identify lack of two-by-two correlation between coverage profile and identify putative outlier sample, distribution of the correlation data is performed. **Here is a summarized boxplot of the two-by-two correlation scores of the entire series**. Putative outlier samples are represented as open circles (Q1 -- 1.5\*IQR).

# Coverage profile: serie comparison

First, to help in identifying distinct coverage profile, coverage profiles are compared between the whole series. **Here is given a Component Analysis (COA) plot that illustrates the distance between each sample based on its coverage profile**. Isolated samples can be considered as putative outliers.

```{r}
plot_coa(fcount_matrix)
```
